---
title: "structure"
author: "Toshikazu, Matsumura"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{structure}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

  # ```{r setup}
```{r, eval = FALSE}
devtools::load_all(".")
  # library(moranajp)
library(tidyverse)
library(grid)

review <- 
  review_sudachi_c %>%
  unescape_utf() %>%
  dplyr::filter(.data[["text_id"]] < 20) %>%
  clean_up() %>%
  # 
  delete_parenthesis() %>%
  align_sentence(s_id = "text_id")

review %>%
  ggplot(aes(x, .data[[s_id]], label = term)) + 
  geom_text() + 
  theme_bw()
```

```{r, eval = FALSE}
review  %>%
  #   dplyr::filter(text_id < 5) %>%
  dplyr::select(term, x) %>%
  dplyr::group_by(term, x) %>%
  tally() %>%
  arrange(desc(n))

grid::grid.newpage()
vp <- viewport(width = max(text$x))
grid::grid.text(text[[surface_form]], text$x, just = c("right", "centre"), vp = vp)

```

```{r, eval = FALSE}
review_mecab   %>% colnames() %>% stringi::stri_unescape_unicode()
review_ginza   %>% colnames() %>% stringi::stri_unescape_unicode()
review_chamame %>% colnames() %>% stringi::stri_unescape_unicode()
```

```{r, eval = FALSE}
width <- function(x, unit = "mm"){
  grid::stringWidth(x) %>%
  grid::convertWidth(unit = unit)
}
one_jp_width <- function(unit = "mm"){
  space <- stringi::stri_unescape_unicode("\u3000") 
  grid::stringWidth(space) %>%
  grid::convertWidth(unit = unit)
}
```
